---
title: "module2"
author: "Hao Qin  "
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE,warning = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.Cleaning data
```{r,echo = FALSE, warning = FALSE, message=FALSE}
rm(list=ls())
library(car)
library(knitr)
library(carData)
library(plyr)
library(stringr)
#setwd(paste0(strsplit(getwd(),split = "/code"), "/dataset"))
bodyfat=read.csv('BodyFat.csv')
#boxplot(bodyfat)
```

## 2.Detection of the outliers and their treatment 

We use three different types of method to detect not only the outliers, but also the influence points in the original dataset, 
which is Siri's equation, BMI formula and cook's distance. Those points will deviate from the line or the region on which most points are, 
thus, it is easy for us to detect them through graphics.
```{r, echo = FALSE, warning = FALSE,fig.height = 2}
table <- data.frame(Outliers = c(39,42,48,76,86,96,163,182,216,221), 
                    Detection_method = c('Cook\'s distance','BMI formula','Siri\'s equation','Siri\'s equation','influence test',
                                         'Siri\'s equation','BMI formula','Siri\'s equation','Siri\'s equation','BMI formula'),
                    Abnormal_Component = c('large Cook\'s distance','too short in height','abnormal body fat','abnormal body fat','deviating from the majority',
                                           'abnormal body fat','abnormal BMI index','zero body fat','density below than 1','lighter than normal'),
                    Treatment = c('deleting','imputation','imputation','imputation','deleting','imputation','imputation','deleting','deleting','imputation'))
#print(table)
kable(table)

layout(matrix(1:4,ncol= 4))
#siri's theatre
reverse.density=1/bodyfat$DENSITY
plot(bodyfat$BODYFAT ~ reverse.density,main="Siri's equation",xlab="1/Density",ylab="Body fat",pch=19,col="red")
text(reverse.density[c(96, 182 ,48 ,76)],bodyfat$BODYFAT[c(96, 182 ,48 ,76)] ,c(96, 182 ,48 ,76),p=4) 
text(reverse.density[216],bodyfat$BODYFAT[216],c(216),p = 2)#96 182 48 76 216
#48th entry
bodyfat$BODYFAT[48]=495/bodyfat$DENSITY[48]- 450
#96th entry
bodyfat$DENSITY[96]=495/(bodyfat$BODYFAT[96]+450)
#deleting 182th entry
#76th entry
bodyfat$BODYFAT[76]=495/bodyfat$DENSITY[76]- 450
#deleting 216th entry


#BMI's theatre
bmi=bodyfat$WEIGHT/(bodyfat$HEIGHT)^2
plot(bodyfat$ADIPOSITY[c(-182,-216)] ~ bmi[c(-182,-216)],main="BMI equation",xlab=expression(WEIGHT/(HEIGHT)^2),ylab="ADIPOSITY",pch=19,col="red")
text(bmi[c(42,163)],bodyfat$ADIPOSITY[c(42,163)]  ,c(42,163),p=3) #42,163,221
text(bmi[c(221)],bodyfat$ADIPOSITY[c(221)]  ,c(221),p=1) #42,163,221
#42th entry
bmi.42=(bodyfat$ADIPOSITY[42]-0.07035)/702.07386
height<-sqrt(bodyfat$WEIGHT[42]/bmi.42)
bodyfat$HEIGHT[42]=height
#211the entry
bmi.221=(bodyfat$ADIPOSITY[221]-0.07035)/702.07386
weight=bmi.221*(bodyfat$HEIGHT[221])^2
bodyfat$WEIGHT[221]<-weight
#163th entry
bmi.163=bodyfat$WEIGHT[163]/(bodyfat$HEIGHT[163])^2
adi=0.1942+698.6151*bmi.163
bodyfat$ADIPOSITY[163]<-adi

#39th entry
model=lm(BODYFAT~ ., data=bodyfat[-c(182,216 ),c(-1,-3)])
plot(model,which = 4)
n=nrow(bodyfat[-c(182,216 ),c(-1,-3)])
p=ncol(bodyfat[-c(182,216 ),c(-1,-3)])
abline(h = 4/(n-p),lty=2,col='red')
#86th entry
model_2=lm(BODYFAT~ ., data=bodyfat[-c(39,182,216 ),c(-1,-3)])
plot(model_2,which = 4)
n=n-1#delete 39
abline(h = 4/(n-p),lty=2,col='red')
```
```{r,echo = FALSE, warning = FALSE, message=FALSE, fig.height = 2}
# rechecking from the plots
layout(matrix(1:3,ncol= 3))
plot(bodyfat$BODYFAT[-c(182,216)] ~ reverse.density[-c(182,216)],
     main="Siri's equation \n without the 216th entry",xlab="1/Density",ylab="Body fat",pch=19,col="red",xlim=c(0.9,1.05),ylim=c(0,50))

bmi=bodyfat$WEIGHT/(bodyfat$HEIGHT)^2
plot(bodyfat$ADIPOSITY ~ bmi,main="BMI equation",xlab=expression(WEIGHT/(HEIGHT)^2),ylab="ADIPOSITY",pch=19,col="red")
text(bmi[c(42,163,221)],bodyfat$ADIPOSITY[c(42,163,221)]  ,c(42,163,221),p=2) #96 182 48 76

model_4=lm(BODYFAT~ ., data=bodyfat[-c(39,86,182,216 ),c(-1,-3)])
plot(model_4,which = 4)
n=n-1#delete 42
abline(h = 4/(n-p),lty=2,col='red')
bodyfat_new=bodyfat[-c(39,86,182,216),c(-1)]
```

## 3.Selecting variables
In this part, several methods have been applied, including stepwise aic and bic selection, lasso and group lasso, mallow’s cp and Bess, which is a new proposed way to selecting variables. After applying these methods, since the total sample size is not large, we can use cross validation to measure the performance of each method, to decide which component should be treated as the independent variable in the final model. To be specific, the final model should be as a rule of thumb, which requires that the quantity of the independent variables should not exceed four.
```{r,echo = FALSE, warning = FALSE, message=FALSE}
#aic and bic
##From the BIC method
CVgroup=function(k,Ndata){
  cvlist=list()
  allseq=rep(1:k,ceiling(Ndata/k))[1:Ndata]   
  temp=sample(allseq,Ndata)   
  x=1:k
  dataseq=1:Ndata
  cvlist=lapply(x,function(x) {dataseq[temp==x]})  
  return(cvlist)
}

mygroup=CVgroup(5,nrow(bodyfat_new))


#calculate the mean mse after cross validation
DoCV=function(k,insertdata,model){
  mygroup=CVgroup(k,nrow(insertdata))
  datacol=ncol(insertdata)
  MSEs=1:k
  for(i in 1:5) {
    Ntraining = setdiff(1:5, i)
    trainingINDEX = numeric()
    
    for (t in Ntraining) {
      trainingINDEX = c(trainingINDEX, unlist(mygroup[[t]]))
    }
    
    testINDEX = mygroup[[i]]
    
    m = lm(model, data  = insertdata[trainingINDEX, ])
    
    test_X = insertdata[testINDEX, 2:datacol]
    test_Y = insertdata[testINDEX, 1]
    
    predicted_Y = predict(m, test_X)
    MSEs[i] = sum((test_Y - predicted_Y)^2) / length(test_Y)
  }
  return(mean(MSEs))
}

#calculate the mse after s times
CVMSE=function(insertdata,model,k,s){
  sumMES = 0
  
  for (i in 1:s) {
    sumMES = sumMES + DoCV(k,insertdata, model)
  }
  
  return(sumMES / s)
}
model.new <- lm(BODYFAT~ WEIGHT + ABDOMEN + WRIST,data=bodyfat_new) # the R^2 is 0.7251
res.bic.4 <- model.new$residuals
model.bic1=BODYFAT ~ WEIGHT + ABDOMEN + WRIST
bic_cv1 = CVMSE(bodyfat_new, model.bic1, 5,100)
#leaving wrist out
model.new1<-lm(BODYFAT~ WEIGHT + ABDOMEN,data=bodyfat_new)
model.bic2=BODYFAT ~ WEIGHT + ABDOMEN
bic_cv2 = CVMSE(bodyfat_new, model.bic2, 5,100)
#one variable, part one
model.new2=lm(BODYFAT~ WEIGHT,data=bodyfat_new)
model.bic3=BODYFAT ~ WEIGHT
bic_cv3 = CVMSE(bodyfat_new, model.bic3, 5,100)
#one variable,part two
model.new3=lm(BODYFAT~  ABDOMEN ,data=bodyfat_new)
model.bic3=BODYFAT ~ ABDOMEN 
bic_cv4 = CVMSE(bodyfat_new, model.bic3, 5,100)
```

```{r,echo = FALSE, warning = FALSE, message=FALSE, fig.height = 2.5, fig.width = 7}
# adjusted R^2 square
library(leaps)
library(faraway)
layout(matrix(1:3,ncol= 3))
cleanbodyfat = bodyfat_new
n=ncol(cleanbodyfat)
X = cleanbodyfat[, 3:n]
Y = cleanbodyfat$BODYFAT
a<-regsubsets(X,Y)
plot(a,scale="adjr2")
g2=leaps(X, Y, nbest = 1,method = "adjr2")
#print(colnames(cleanbodyfat)[as.numeric(which(g2$which[9,]==TRUE)) + 3])
model_adjr20 = BODYFAT ~ AGE + WEIGHT + NECK + ABDOMEN +  HIP + THIGH + BICEPS + FOREARM + WRIST
adjr20_cv = CVMSE(cleanbodyfat, model_adjr20, 5,100)   # number of variables is 9
#print(colnames(cleanbodyfat)[as.numeric(which(g2$which[8,]==TRUE)) + 3])
model_adjr21 = BODYFAT ~ AGE + WEIGHT + NECK + ABDOMEN + THIGH + BICEPS + FOREARM + WRIST
adjr21_cv = CVMSE(cleanbodyfat, model_adjr21, 5,100)   # number of variables is 8
#print(colnames(cleanbodyfat)[as.numeric(which(g2$which[7,]==TRUE)) + 3])
model_adjr22 = BODYFAT ~ AGE + WEIGHT + NECK + ABDOMEN + THIGH  + FOREARM + WRIST
adjr22_cv = CVMSE(cleanbodyfat, model_adjr22, 5,100)   # number of variables is 7
#print(colnames(cleanbodyfat)[as.numeric(which(g2$which[6,]==TRUE)) + 3])
model_adjr23 = BODYFAT ~ AGE + WEIGHT + ABDOMEN + THIGH  + FOREARM + WRIST
adjr23_cv = CVMSE(cleanbodyfat, model_adjr23, 5,100)   # number of variables is 6
#print(colnames(cleanbodyfat)[as.numeric(which(g2$which[5,]==TRUE)) + 3])
model_adjr24 = BODYFAT ~ AGE + WEIGHT + ABDOMEN + THIGH + WRIST
adjr24_cv = CVMSE(cleanbodyfat, model_adjr24, 5,100)   # number of variables is 5
#print(colnames(cleanbodyfat)[as.numeric(which(g2$which[4,]==TRUE)) + 3])
model_adjr25 = BODYFAT ~ WEIGHT + ABDOMEN + BICEPS + WRIST
adjr25_cv = CVMSE(cleanbodyfat, model_adjr25, 5,100)   # number of variables is 4
#print(colnames(cleanbodyfat)[as.numeric(which(g2$which[3,]==TRUE)) + 3])
model_adjr26 = BODYFAT ~ WEIGHT + ABDOMEN +  WRIST
adjr26_cv = CVMSE(cleanbodyfat, model_adjr26, 5,100)   # number of variables is 3
#print(colnames(cleanbodyfat)[as.numeric(which(g2$which[2,]==TRUE)) + 3])
model_adjr27 = BODYFAT ~ WEIGHT + ABDOMEN
adjr27_cv = CVMSE(cleanbodyfat, model_adjr27, 5,100)   # number of variables is 2
#print(colnames(cleanbodyfat)[as.numeric(which(g2$which[1,]==TRUE)) + 3])
model_adjr28 = BODYFAT ~ ABDOMEN 
adjr28_cv = CVMSE(cleanbodyfat, model_adjr28, 5,100)   # number of variables is 1
adjr2_cv=c(adjr20_cv,adjr21_cv,adjr22_cv,adjr23_cv,adjr24_cv,adjr25_cv,adjr26_cv,adjr27_cv,adjr28_cv)
plot(x=9:1,adjr2_cv,xlab = "number of variavbles",ylab="Mean square error",type = "b",main = "MSE vs # of variables")
plot(x=1:9,g2$adjr2[1:9],xlab = "number of variavbles",ylab="Adjusted R Square",type = "b",main = "Adjusted R Square vs # of variables")
```



```{r,echo = FALSE, warning = FALSE, message=FALSE, fig.height = 2.5, fig.width = 7}
#mallow's cp
layout(matrix(1:3,ncol= 3))
cleanbodyfat = bodyfat_new
n=ncol(cleanbodyfat)
X = cleanbodyfat[, 3:n]
Y = cleanbodyfat$BODYFAT
a=regsubsets(X,Y)
plot(a,scale="Cp")
g1=leaps(X, Y, nbest = 1,method = "Cp")
#Cpplot(g1)
#order(g1$Cp)[1] #7
model_cp0 = BODYFAT ~ AGE + WEIGHT + NECK + ABDOMEN  + THIGH + BICEPS + FOREARM + WRIST
cp0_cv = CVMSE(cleanbodyfat, model_cp0, 5,100)   # number of variables is 8
#print(colnames(cleanbodyfat)[as.numeric(which(g1$which[7,]==TRUE)) + 3])
model_cp1 = BODYFAT  ~ AGE + WEIGHT + NECK + ABDOMEN  + THIGH + FOREARM + WRIST
cp1_cv = CVMSE(cleanbodyfat, model_cp1, 5,100)   # number of variables is 7
#print(colnames(cleanbodyfat)[as.numeric(which(g1$which[6,]==TRUE)) + 3])  
model_cp2 = BODYFAT  ~ AGE + WEIGHT + ABDOMEN + THIGH + FOREARM + WRIST
cp2_cv = CVMSE(cleanbodyfat, model_cp2, 5,100)   # number of variables is 6
#print(colnames(cleanbodyfat)[as.numeric(which(g1$which[5,]==TRUE)) + 3])  
model_cp3 = BODYFAT  ~  WEIGHT + NECK + ABDOMEN + THIGH + WRIST
cp3_cv = CVMSE(cleanbodyfat, model_cp3, 5,100)   # number of variables is 5
#print(colnames(cleanbodyfat)[as.numeric(which(g1$which[4,]==TRUE)) + 3])  
model_cp4 = BODYFAT  ~  WEIGHT + ABDOMEN + BICEPS + WRIST
cp4_cv = CVMSE(cleanbodyfat, model_cp4, 5,100)   # number of variables is 4
#print(colnames(cleanbodyfat)[as.numeric(which(g1$which[3,]==TRUE)) + 3])  
model_cp5 = BODYFAT  ~  WEIGHT + ABDOMEN + WRIST
cp5_cv = CVMSE(cleanbodyfat, model_cp5, 5,100)   # number of variables is 3
#print(colnames(cleanbodyfat)[as.numeric(which(g1$which[2,]==TRUE)) + 3])  
model_cp6 = BODYFAT  ~  WEIGHT + ABDOMEN 
cp6_cv = CVMSE(cleanbodyfat, model_cp6, 5,100)   # number of variables is 2
#print(colnames(cleanbodyfat)[as.numeric(which(g1$which[1,]==TRUE)) + 3])  
model_cp7 = BODYFAT  ~ ABDOMEN 
cp7_cv = CVMSE(cleanbodyfat, model_cp7, 5,100)   # number of variables is 1
cp_cv=c(cp0_cv,cp1_cv,cp2_cv,cp3_cv,cp4_cv,cp5_cv,cp6_cv,cp7_cv)
#par(mfrow=c(1,2))
plot(x=8:1,cp_cv,xlab = "number of variavbles",ylab="Mean square error",type = "b",main = "MSE vs # of variables")
plot(x=1:8,y=g1$Cp[1:8],xlab = "number of variavbles",ylab="Mallow's cp",type = "b",main = "Cp vs # of variables")
```

```{r}
#lasso and group lasso
```

```{r,echo = FALSE, warning = FALSE, message=FALSE, fig.height = 2, fig.width = 7}
library("BeSS")
total=bodyfat_new
total = total[,-2]
aic_bess = c()
bic_bess = c()
gic_bess = c()
mse_bess = c()
aic_bess[1] = Inf
bic_bess[1] = Inf
gic_bess[1] = Inf
mse_bess[1] = Inf
for ( i in 2:(length(total[1,])-1)){
  model_bic <- bess.one(x = total[,-1],y = total[,1],s = i)
  mse_bess[i] = model_bic$mse
  aic_bess[i] = model_bic$AIC
  bic_bess[i] = model_bic$BIC
  gic_bess[i] = model_bic$GIC
}
#c(which(mse_bess[] == min(mse_bess)), which(aic_bess[] == min(aic_bess)),which(bic_bess[] == min(bic_bess)),which(gic_bess[] == min(gic_bess)))
layout(matrix(1:4,ncol = 4))
plot(x = 2:length(total[1,])-1, y = mse_bess,xlab = "number of variables",ylab = "mse")
plot(x = 2:length(total[1,])-1, y =aic_bess,xlab = "number of variables",ylab = "AIC")
plot(x = 2:length(total[1,])-1, y =bic_bess,xlab = "number of variables",ylab = "BIC")
plot(x = 2:length(total[1,])-1, y =gic_bess,xlab = "number of variables",ylab = "GIC")
model_bess_5 <- bess.one(y = total[,1], x = total[,-1],s = 5)
model_bess_4 <- bess.one(y = total[,1], x = total[,-1],s = 4)
model_bess_3 <- bess.one(y = total[,1], x = total[,-1],s = 3)
#CV
times = 1000
mse_bess_5 <- rep(0,times)
mse_bess_4 <- rep(0,times)
mse_bess_3 <- rep(0,times)
total_bess_5 <- rep(0,times)
total_bess_4 <- rep(0,times)
total_bess_3 <- rep(0,times)
for (g in 1:times){
  choices = sample(1: length(total[,1]),82*2)
  test <- total[as.vector(choices),]
  val <- total[-as.vector(choices),]
  
  model_bess_3 <- bess.one(y = test[,1], x = test[,c(4,7,8)],s = 3)
  pre_bess_3 <-predict(model_bess_3,val,type = "BIC")
  
  model_bess_5 <- bess.one(y = test[,1], x = test[,c(3,5,7,8,15)],s = 5)
  pre_bess_5 <-predict(model_bess_5,val,type = "BIC")
  
  model_bess_4 <- bess.one(y = test[,1], x = test[,c(3,5,7,8)],s = 4)
  pre_bess_4 <-predict(model_bess_4,val,type = "BIC")
    
  ave <- ave(val[,1])
  for( i in 1:length(val[,1])){
    mse_bess_5[g] = as.numeric((pre_bess_5[i] - val[i,1])^2) + mse_bess_5[g]
    mse_bess_4[g] = as.numeric((pre_bess_4[i] - val[i,1])^2) + mse_bess_4[g]
    mse_bess_3[g] = as.numeric((pre_bess_3[i] - val[i,1])^2) + mse_bess_3[g]
    total_bess_5[g] = as.numeric((val[i,1] - ave[i])^2) + total_bess_5[g]
    total_bess_4[g] = as.numeric((val[i,1] - ave[i])^2) + total_bess_4[g]
    total_bess_3[g] = as.numeric((val[i,1] - ave[i])^2) + total_bess_3[g]
  }
}
R2_5 = mean(1 - mse_bess_5/total_bess_5)
R2_4 = mean(1 - mse_bess_4/total_bess_4)
R2_3 = mean(1 - mse_bess_3/total_bess_3)
mse_bess_5_cv = mean(mse_bess_5/(length(total[,1])-82*2))
mse_bess_4_cv = mean(mse_bess_4/(length(total[,1])-82*2))
mse_bess_3_cv = mean(mse_bess_3/(length(total[,1])-82*2))
```


```{r,echo = FALSE, warning = FALSE, message=FALSE}
table2 <- data.frame(model = c(rep('BIC', 4), rep('Mallow\'s Cp', 4), rep('Adjusted R square', 4), rep('BeSS', 3)),
                     Mean_Sq = c(bic_cv1, bic_cv2, bic_cv3, bic_cv4, 
                                 cp7_cv, cp6_cv, cp5_cv, cp4_cv, 
                                 adjr28_cv, adjr27_cv, adjr26_cv, adjr25_cv,
                                 mse_bess_3_cv, mse_bess_4_cv, mse_bess_5_cv),
                     variables = c('WEIGHT + ABDOMEN + WRIST', 'WEIGHT + ABDOMEN', 'WEIGHT', 'ABDOMEN', 
                                   'ABDOMEN', 'WEIGHT + ABDOMEN', 'WEIGHT + ABDOMEN + WRIST', 'WEIGHT + ABDOMEN + BICEPS + WRIST', 
                                   'ABDOMEN', 'WEIGHT + ABDOMEN', 'WEIGHT + ABDOMEN +  WRIST', 'WEIGHT + ABDOMEN + BICEPS + WRIST',
                                   'HEIGHT + CHEST + ABDOMEN', 'ADIPOSITY + CHEST + ABDOMEN + HIP ', 'WEIGHT + ADIPOSITY + CHEST + ABDOMEN + WRIST'
                                   ))
table2  = table2[order(table2$Mean_Sq),]
kable(table2)
```

## Model building